---
title: "FE vs RE"
author: Nelson Auner
date: May 6 2015
output: md_document
---

Test for whether you should use fixed or random effects
========================================================

I've been thinking about random effects this week. I don't think many people in business use them and I'm not sure if there's a good reason for this. One explanation is that fixed effects are common in introductory statistics and econometrics classes, so many people doing "data" work have not heard of them. A second explanation might be that the craze around machine learning has encouraged people to hack at data without much regard to the underlying statistical processes.  

My thought here is to try both on a couple of different data sets to understand the difference!

## 1 - Corn Sites
from http://www.stat.wisc.edu/courses/st572-larget/Spring2007/handouts13-2.pdf

Random effect assumes

- $y_{ij} = \mu + a_i + e_{ij}, j=1,...,n, i=1,...,k$

- $a_i \sim N(0,\sigma_a^2)$

- $e_i \sim N(0,\sigma_e^2)$



```{r setup, message=FALSE,warning=FALSE,results='hide'}
# thanks dirk
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

#install.packages("dplyr")
library(dplyr)
#install.packages("lme4")
library(lme4)

#install.packages("DAAG")
library(DAAG)


```

So, here is what the data looks like. We observe *harvwt* clustered by site, also some variance components...

```{r ant111b}
str(ant111b)
xyplot(harvwt ~ site, pch = 16, col = "blue",data=ant111b)
# still not clear why FE is a bad idea
corn1.lmer = lmer(harvwt ~ 1 + (1 | site),data=ant111b)
summary(corn1.lmer)
```
Always check residuals!
```{r ant11b_residuals}
plot(fitted(corn1.lmer), residuals(corn1.lmer))
```
So here's one difference--the fitted values are not the sample means (as in FE), but rather shrinkage estimates. 

```{r ant11b_shrinkage}
with(ant111b, {
  means = sapply(split(harvwt, site), mean)
  siteFit = sapply(split(fitted(corn1.lmer), site), mean)
  data.frame(mean = means, fitted = siteFit)
})
# means
# returns error since 'with' command keeps objects in temporary namespace (pretty cool)
with(ant111b, {
corn1.aov = aov(harvwt ~ Error(site))
summary(corn1.aov)
})
```

