The popularity of `dplyr` has led to the use of `>%>` and an increase in the awareness of functional programming in R. This is great, as side-effectful statements has long been one of my greatest causes of error. Errors are scary, and it's halloween, so here's a quick example of `>%>` for `dplyr` and `ggplot`, using M&M data from the University of Puget Sound's [Data Hoard](http://stat.pugetsound.edu/hoard/datasets.aspx). Although the  [dev version](https://github.com/smbache/magrittr)  includes some new operators, this post covers only the standard `%>%` operator, which can be summarized as ` x %>% f == f(x)`


First, we load the data and produce a quick summary table by type of m&m. The `>%>` operator isn't necessary here but helps keep the code structured and readable. 

```{r loaddata}
library(reshape2) # http://bit.ly/1x2OEv5
library(magrittr) # ceci n'est pas un pipe
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
mms <- read.csv('http://stat.pugetsound.edu/hoard/datasets/mms.csv')
mms %>% summary

mms %>% 
  group_by(type) %>% 
  summarise(m_diam = mean(diameter), m_mass = mean(mass))

mms %>% melt %>% head # equivalent to head(melt(mms)) 
```

We also might want to investigate size variation by both type and color: 

```{r dplyr}
mms  %>%  
  group_by(type,color) %>% 
  summarise(d=mean(diameter),m=mean(mass),count = n())
```

There seems to be more variation in the peanut m&ms. We make a spooky graph to investigate. 
Note that `>%>` works fine inside other functions, and helps keep the code uncluttered from parenthesis.


```{r ggplot,fig.width=8, fig.height=6}
ggplot(mms,
       aes(x=diameter,
           y=mass,
           size=type,
           colour=color)) +  
  scale_color_manual(values = 
                       mms$color %>% 
                       unique %>% 
                       as.character %>% sort) +
  scale_size_manual(values = c(6,4,2)) + 
  geom_point(alpha=2/3) + 
  stat_smooth(method=loess,fill="orange",alpha=.5,colour="orange",size=1,aes(group=type)) + 
  ggtitle("Mass vs. Diameter of M&Ms") + 
  theme(legend.key = element_rect(colour = "black")) +
   theme(panel.background = element_rect(fill = "black"))


```



References: 

[ggplot and dplyr at timelyportfolio](http://timelyportfolio.blogspot.com/2014/07/alternate-price-plots-ggplot2-magrittr.html)

[ggplot2 layers](http://rstudio-pubs-static.s3.amazonaws.com/3355_d3f08cb2f71f44f2bbec8b52f0e5b5e7.html)

[maggritr on github](https://github.com/smbache/magrittr)  
  
  
